---
title: "web_scraping"
output: html_document
date: "2023-01-09"
---

# API call 

*Zuwendungsdatenbank*
In der Zuwendungsdatenbank werden die von den Behörden bewilligten Zuwendungen an juristische Personen (ab einem Zuwendungsbetrag i. H. v. 100 Euro) rückwirkend für die letzten 5 Jahre veröffentlicht. Dies erfolgt jeweils zum 30. Juni eines jeden Jahres. Aktuell enthält die Datenbank die gewährten Zuwendungen der Jahre 2017-2021. 

```{r}
library(jsonlite)
```

https://www.berlin.de/sen/finanzen/service/zuwendungsdatenbank/index.php/api.html 

Application Programming Interface 
- Damit apps / programme miteinander sprechen können. 
- Wenn du ein Uber buchst: 100+ apps die angesprochen werden, bsw. für das Routing oder Bezahlung
- APIs sind das User Interface von Code:
https://www.berlin.de/sen/finanzen/service/zuwendungsdatenbank/


Restricted access
- API Keys --> Damit wir individuell zuordnenbar sind und der API weiß wer wir sind
- Sonst können wir einen Server overloaden (und wie eine DOS attacke ausschauen) 
- Falls es keinen Key gibt können wir bsw. sys

## Defining the API

- Im Anwendungsbereich ist in unserem Fall ein API-call im Grunde einfach eine URL 
- In der URL sind bestimmte Informationen gespeichert. 
- Welche Informationen das sind finden wir in der Dokuemntation 

```{r}
# api documentation
# https://www.berlin.de/sen/finanzen/service/zuwendungsdatenbank/
  
# base url 
base_url <- "https://www.berlin.de/sen/finanzen/service/zuwendungsdatenbank/index.php"

# who financed the project?
geber = "Bezirksamt+Neukölln"
# which data format?
data = ".json"

# we generate the url
neukoelln <- paste0(base_url,
                    "/.json?q=",
                    "&geber=",geber)

# and grab the json using the package jsonlite
df_nk <- fromJSON(neukoelln)
```

lets look at what we got 
```{r}
df_nk$messages
df_nk$results
df_nk$index
df_nk$item
```

We only get the results from one page! 
--> how many pages do we have to scrape?
```{r}
# how many total results?
df_nk$results$count
# how many per page?
df_nk$results$items_per_page
# so how many pages to scrape?
n_pages <- df_nk$results$count/df_nk$results$items_per_page
# we need to round to the next larger integer 
n_pages_ceiling <- ceiling(n_pages)
```

we add a page number to our api call 
```{r}
page_number <- 1
neukoelln_pages <- paste0(neukoelln, 
                          "&page=", 
                          page_number)
df_nk_pages <- fromJSON(neukoelln_pages)
df_nk_pages$index

page_number <- 2
neukoelln_pages <- paste0(neukoelln, 
                          "&page=", 
                          page_number)
df_nk_pages <- fromJSON(neukoelln_pages)
df_nk_pages$index
```

and now we need to loop through all the pages
```{r loop: a test run}
#for (i in n_pages_ceiling) {
for (i in 1:2) {
  page_number = i
  neukoelln_pages <- paste0(neukoelln, 
                          "&page=", 
                          page_number)
  df_nk_pages <- fromJSON(neukoelln_pages)
  print(df_nk_pages$index)
}
```
combine the results to one df 
```{r loop: a test run}
# create a base df that will be populated in the loop
base_df <- df_nk$index[0,]
#for (i in n_pages_ceiling) {
for (i in 1:2) {
  page_number = i
  print(paste0("Scraper is on page ", i))
  neukoelln_pages <- paste0(neukoelln, 
                          "&page=", 
                          page_number)
  df_nk_pages <- fromJSON(neukoelln_pages)
  base_df <- rbind(base_df, 
                   df_nk_pages$index)
}
```

## Create the final scraper 
```{r final scraper}
# create a base df that will be populated in the loop
base_df <- df_nk$index[0,]
#for (i in n_pages_ceiling) {
for (i in 1:n_pages_ceiling) {
  # clear the output 
  cat("\014")      
  # define the page number 
  page_number = i   
  # print our progress
  print(paste0("Scraper is on page ", i, " of ",n_pages_ceiling))
  # call the API
  neukoelln_pages <- paste0(neukoelln, 
                          "&page=", 
                          page_number)
  # create a df from the api call 
  df_nk_pages <- fromJSON(neukoelln_pages)
  # append it to our existing df
  base_df <- rbind(base_df, 
                   df_nk_pages$index)
}
```

next step: define a function for different geber 
```{r fincal scraper function}
finanz_scraper <- function(geber){
  
  #### we generate a first api call
  
  # base url 
  base_url <- "https://www.berlin.de/sen/finanzen/service/zuwendungsdatenbank/"
  # who financed the project? --> the argument of our function! 
  geber = geber
  # which data format?
  data = ".json"
  
  # we generate the url
  finance_url <- paste0(base_url,
                      "index.php/index/index.json?q=",
                      "&geber=",geber)
  
  # and do our first api call 
  finance_df <- fromJSON(finance_url)

  #### we look at the results 
  
  # how many total results?
  finance_df$results$count
  # how many per page?
  finance_df$results$items_per_page
  # so how many pages to scrape?
  n_pages <- finance_df$results$count/finance_df$results$items_per_page
  # we need to round to the next larger integer 
  n_pages_ceiling <- ceiling(n_pages)
  
  #### we call the API 
  base_df <- finance_df$index[0,]
  #for (i in n_pages_ceiling) {
  for (i in 1:n_pages_ceiling) {
    # clear the output 
    cat("\014")      
    # define the page number 
    page_number = i   
    # print our progress
    print(paste0("Geber: ",geber, " | Scraper is on page ", i, " of ",n_pages_ceiling))
    # call the API
    finance_pages <- paste0(finance_url, 
                            "&page=", 
                            page_number)
    # create a df from the api call 
    df_finance_temporary <- fromJSON(finance_pages)
    # append it to our existing df
    base_df <- rbind(base_df, 
                     # we access the df of the api call 
                     df_finance_temporary$index)
  }
  return(base_df)
}
```

call the function
```{r}
finanz_scraper(geber = paste("Bezirksamt+Mitte"))
```

scape n bezirke
```{r}
districts <- c("Mitte", "Neukölln", "Treptow-Köpenick")

# create empty df
final_df=NULL

# call function in loop
for (district in districts) {
  temp_df <- finanz_scraper(geber = paste0("Bezirksamt+", district))
  # populate df
  final_df <- rbind(final_df, 
                    temp_df)
}
```

# Analysis

```{r libraries}
library(dplyr)
library(ggplot2)
```


```{r}
# financing by year by district 
financing_jahr <- final_df %>%
  group_by(geber, jahr) %>%
  summarise(gesamtbetrag = sum(as.numeric(betrag)))

ggplot(financing_jahr, aes(x=as.numeric(jahr), 
                           y=gesamtbetrag, 
                           col=geber)) +
  geom_line() +
  theme_bw()

# average zuwendung?
ggplot(final_df, aes(x=log(as.numeric(betrag)), 
                     fill=geber)) +
  geom_density(alpha=0.5) +
  theme_bw()
```

